:::::::: {.content-cell data-scenario="bau" data-subject="econ"}

:::::: columns
::: {.column width="50%"}
Ecosystem services influence a wide range of economic sectors, and future changes in these services will affect the Peruvian economy. To assess this impact, we examined the relationships between **8 ecosystem services** and **7 economic sectors**, using data from the [ENCORE database](https://www.encorenature.org/en), which evaluates sectoral dependencies on ecosystem services. Only the sectors relevant to Peru were included.

We analyzed **two scenarios** to evaluate the impact of ecosystem service changes on economic sectors. We calculate the impact as the product of the **average change** in ecosystem services and the **materiality rating** of each economic sector: 

$$
Impact = avg\_change \times mat\_rating + std
$$
The results of the scenario analyses are shown below in the two bar plots. 
:::

::: {.column width="50%"}
<div style="font-size:12px; width:80%;">

```{r}
#| echo: false
#| warning: false
library(knitr)
library(kableExtra)

# Load data
scenarios <- read.csv("outputs/data/dummy_scenarios.csv")

# Smaller table
kable(scenarios, format = "html") %>%
  kable_styling(
    full_width = FALSE,
    font_size = 10,
    bootstrap_options = c("striped", "hover", "condensed")
  )
```
</div>
:::


::: {.column width="50%"}
```{r}
#| echo: false
#| warning: false
#| fig-width: 8
#| fig-height: 8
library(dplyr)
library(tidyr)
library(plotly)
# Load ENCORE ES and economic impact for Peru DF from the data folder
peru <- read.csv("outputs/data/peru_diagram_df.csv")

peru$ISIC_cat[peru$ISIC_cat=="Public_admin"] <- "Public Administration"

#load scenario df
scenarios <- read.csv("outputs/data/dummy_scenarios.csv")

# map rating to numeric scores 
rating_map <- c("VL" = 1, "L" = 2, "M" = 3, "H" = 4, "VH" = 5)

peru <- peru %>%
  mutate(rating_num = ifelse(is.na(rating), 0, rating_map[as.character(rating)]))

# join peru df with scenarios 
scenario_data <- peru %>%
  inner_join(scenarios, by = "Ecosystem_Service")

# expand each row of peru_df for all scenarios
#expanded <- crossing(Scenario = unique(scenarios$Scenario), peru)
# now join safely
#joined <- expanded %>%
#  left_join(scenarios, by = c("Scenario", "Ecosystem_Service"))


# calculate impacts with noise 
set.seed(123) # reproducible

scenario_data <- scenario_data %>%
  rowwise() %>%
  mutate(
    impact = ifelse(
      is.na(avg_change),
      0,  # no change if ES not in scenario
      avg_change * rating_num + rnorm(1, mean = 0, sd = std)
    )
  ) %>%
  ungroup()

# box plot 
# Aggregate to ISIC_section (or ISIC_division if you prefer)
category_summary <- scenario_data %>%
  group_by(ISIC_cat, Scenario) %>%
  summarise(
    mean_impact = mean(impact, na.rm = TRUE),
    sd_impact = sd(impact, na.rm = TRUE),
    .groups = "drop"
  )


# Define bins and colors
bin_colors <- c(
  "[-1,-0.75)"  = "#873426",  # dark red
  "[-0.75,-0.5)"= "#A64B2A",  # red-orange
  "[-0.5,-0.25)"= "#D27E3F",  # orange
  "[-0.25,0)"   = "#E5C965",  # yellow
  "[0,0.25)"    = "#A8D08D",  # light green
  "[0.25,0.5)"  = "#70AD47",  # medium green
  "[0.5,0.75)"  = "#4F7F3F",  # dark green
  "[0.75,1]"    = "#3B5C3E"   # forest green
)

# Get all ISIC categories (so missing ones are kept in plots)
all_cats <- unique(scenario_data$ISIC_cat)

# Loop over scenarios and store plots in a list
plots <- list()
scenarios_list <- unique(scenario_data$Scenario)

for (scen in scenarios_list) {
  
  scen_data <- scenario_data %>%
    filter(Scenario == scen) %>%
    group_by(ISIC_cat) %>%
    summarise(
      mean_impact = mean(impact, na.rm = TRUE),
      sd_impact   = sd(impact, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    right_join(tibble(ISIC_cat = all_cats), by = "ISIC_cat") %>%
    mutate(
      mean_impact = replace_na(mean_impact, 0),
      sd_impact   = replace_na(sd_impact, 0),
      impact_bin  = cut(mean_impact,
                        breaks = seq(-1, 1, by = 0.25),
                        include.lowest = TRUE,
                        right = FALSE)
    )
  
  # Static ggplot
  p <- ggplot(scen_data, aes(x = ISIC_cat, y = mean_impact, 
                             fill = impact_bin,
                             text = paste0("ISIC: ", ISIC_cat,
                                           "<br>Mean Impact: ", round(mean_impact, 3),
                                           "<br>SD: ", round(sd_impact, 3)))) +
    geom_col(width = 0.7) +
    geom_errorbar(aes(ymin = mean_impact - sd_impact,
                      ymax = mean_impact + sd_impact),
                  width = 0.2) +
    scale_fill_manual(values = bin_colors, drop = FALSE, na.value = "grey90") +
    coord_cartesian(ylim = c(-1, 1)) +
    theme_minimal(base_size = 12) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    labs(
      title = paste(scen),
      x = "ISIC Category",
      y = "Economical Impact",
      fill = "Impact"
    )
  
  # Convert to plotly
  plots[[scen]] <- ggplotly(p, tooltip = "text")
}

# --- Call plots interactively one by one ---
plots[["Scenario 1"]]



```
:::

::: {.column width="50%"}
```{r}
#| echo: false
#| warning: false
#| fig-width: 8
#| fig-height: 8
plots[["Scenario 2"]]
```
:::

